Article Title

### 2013-12-23 plos.tex
rev31 - update of Discussion after conversation with Kreso.

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.  Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.  Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.

@@

We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}). At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D). However, the corresponding partial correlation matrices differed dramatically (Fig.~\ref{fig:05}\,B and E). The partial correlation was decomposed into sparse and low-rank components (Fig.\,\ref{fig:05}\,C). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations. The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G). The low-rank component had rank 17.

@@

The average partial correlations revealed by the sparse+latent estimator at all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (\figref{06}{A}). The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{B and C}). However, there was an inverse relationship between the number of latent units and the average node degree (\figref{fig:06}{D}). Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.  These differences have multiple explanations, including differences in brain states and recording quality (neuropil contamination, motion).

@@

In contrast to functional architecture, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or trial-to-trial variations,  not directly driven by external stimuli.  Functional connectivity is thought to reflect internal, recurrent connectivity in neural circuits.  This line of inquiry ascribes greater significance to the brain's intrinsic activity (often dismissed as `noise') \cite{Yuste:2005}. However, relatively little progress has been made in understanding the role of functional connectivity and its relationship to the anatomical connectivity.

Neural correlations boast a rich history as indicators of functional connectivity. When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970,Alonso:1998,Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many synaptic connections may not produce such signatures). Without targeted stimulation of specific cells or measurements of subthreshold intracellular events, such analysis remains the best method for detecting probable monosynaptic connections in vivo. When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent features of activity with little correspondence to the local anatomical connectivity. These slower correlations correspond more closely to the computation performed by the circuit than to its anatomical organization. For this reason, firing rate correlations at scales of tens to hundreds milliseconds have been studied for their effect on coding under the assumption of rate codes and linear downstream decoder \cite{Averbeck:2006} with little interest in anatomy.  Calcium imaging allows simultaneous measurements of the activity of large populations of cells \emph{in vivo} \cite{Katona:2012,Cotton:2013} and even of entire nervous systems \cite{Leung:2013,Ahrens:2013}, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes, limiting its use to slower temporal scales at which correspondence between functional and anatomical connectivity is less direct.

Our study is part of the effort to find descriptions of functional connectivity with the greatest chance of correspondence to anatomical, causal organization of the circuit.

@@

The MATLAB code for these computations is made publically available at {\tt http://github.com/atlab/cov-est}.

For plots in \figref{fig:05}{B and E}, \figref{fig:06}{A}, and \figref{fig:07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}). The sparse and low-rank components in \figref{fig;05}{C} are the additive components of these normalized partial correlations.

The sparse partial correlations 
\paragraph{Simulation}

>>>>

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.  Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.  Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate. Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.

@@

We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}). At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (\figref{05}{A and D}). However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}). The partial correlation was decomposed into sparse and low-rank components (\figref{ 05}{C}). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations. The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}). The low-rank component had rank 17.

@@

The average partial correlations revealed by the sparse+latent estimator at all 31 sites were about 5 times lower than the sample correlations and  less variable across sites (\figref{06}{A}). The average node degree of the sparse component of the partial correlations and the number of inferred latent units varied widely between sites but generally increased with recorded population size (\figref{06}{B and C}). However, there was an inverse relationship between the number of latent units and the average node degree (\figref{06}{D}). Several sites, even with relatively large population sizes, had fairly few pairwise interactions and were dominated by latent units.  These differences have multiple explanations, including differences in brain states and recording quality (neuropil contamination, motion).

@@

In contrast to functional architecture, \emph{functional connectivity} commonly refers to the degree of correlation, synchronization, or other forms of statistical association between neuronal populations or individual neurons inferred from spontaneous activity or trial-to-trial variations,  not directly driven by external stimuli.  Functional connectivity is thought to reflect internal, recurrent connectivity in neural circuits.   However, less experimental evidence has been garnered to for links between functional and anatomical connectivity.

Neural correlations boast a rich history as indicators of functional connectivity. When measured on temporal scales of milliseconds, sharp peaks or troughs in the cross-correlogram between pairs of neurons at latencies typical of synaptic transmission provide evidence of monosynaptic connection between pairs of cells \cite{Moore:1970, Alonso:1998, Denman:2013} (although other processes can also generate sharp peaks in the cross-correlogram while many monosynaptic connections may not produce such signatures). Without targeted stimulation of specific cells or measurements of subthreshold intracellular events, such analysis remains the best method for detecting probable monosynaptic connections in vivo.

When measured at longer temporal scales, neural correlations reflect activations involving large ensembles of neurons and may reflect direct causal anatomical interactions, indirect causal interactions, or complex network-wide emergent features of activity with little correspondence to the local anatomical connectivity. Calcium imaging allows simultaneous measurements of the activity of large populations of cells \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even of entire nervous systems \cite{Leung:2013, Ahrens:2013}, but its temporal bandwidth is limited to the scale of tens or hundreds milliseconds by the kinetics of calcium-sensitive dyes, limiting its use to slower temporal scales at which correspondence between functional and anatomical connectivity is less direct.

@@

For plots in \figref{05}{B and E}, \figref{06}{A}, and \figref{07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}). The sparse and low-rank components in \figref{05}{C} are the additive components of these normalized partial correlations.

The sparse partial correlations computed for the display of connectivity in \figref{05}{F, G, H}, was computed by normalizing only the sparse component of the precision matrix still using Eq.~\ref{eq:partial}.

The MATLAB code for these computations is available at {\tt http://github.com/atlab/cov-est}.

\paragraph{Simulation}
